Estimating SARS-CoV-2 transmission in England from randomly sampled prevalence surveys

### 2024-03-30 incidence-prevalence.tex
Fix derivation of approximation

Starting with the variances, by the law of total variance:
\begin{align}
  &\var\left( Y_t \mid \vec{Z} \right) \\
    &= \var\left[\E\left( Y_t \mid \vec{Z}, \vec{P} \right) \mid \vec{Z} \right] + \E\left[\var\left( Y_t \mid \vec{Z}, \vec{P} \right) \mid \vec{Z} \right] \\
    &= \var\left( n_t \frac{P_t}{\Npop} \mid \vec{Z} \right) + \E\left( n_t \frac{P_t}{\Npop} \left(1 - \frac{P_t}{\Npop} \right) \mid \vec{Z} \right) &Y_t \mid \vec{P} \dist \Bin \\
    &= \left( \frac{n_t}{\Npop} \right)^2 \var\left(P_t \mid \vec{Z} \right) + \frac{n_t}{\Npop} \E\left(P_t \mid \vec{Z} \right)  - \frac{n_t}{\Npop^2} \E\left(P_t^2 \mid \vec{Z} \right) \\
    &= \left( \frac{n_t}{\Npop} \right)^2 \var\left(P_t \mid \vec{Z} \right) + \frac{n_t}{\Npop} \E\left(P_t \mid \vec{Z} \right)  - \frac{n_t}{\Npop^2} \left(\E\left(P_t \mid \vec{Z} \right) ^ 2 + \V(P_t \mid \vec{X})\right) \\
    &= \frac{n_t(n_t - 1)}{\Npop^2} \var\left(P_t \mid \vec{Z} \right) + n_t \frac{\E\left(P_t \mid \vec{Z} \right)}{\Npop}\left(1 - \frac{\E\left(P_t \mid \vec{Z} \right)}{\Npop} \right) \\
\end{align}
In the contexts I consider $n_t << \Npop$ because $\Npop$ is the population of England (around 56 million) while $n_t$ is the daily sample size (10s of thousands).
Therefore:
\begin{align}
    \frac{n_t(n_t - 1)}{\Npop^2} \var\left(P_t \mid \vec{Z} \right)
    &<< \frac{n_t}{\Npop} \var\left(P_t \mid \vec{Z} \right) \\
    &\leq \frac{n_t}{\Npop} \E\left(P_t \mid \vec{Z} \right) &\text{by \cref{inc-prev:eq:boundVPt}} \\ 
    &\leq \frac{n_t}{\Npop} \E\left(P_t \mid \vec{Z} \right) \left(1 - \frac{\E(P_t \mid \vec{Z}}{\Npop} \right) &\text{as $P_t \leq \Npop$.}
\end{align}
Combining these results:
\begin{align}
  \var\left( Y_t \mid \vec{Z} \right)
    &\approx \frac{\E\left(P_t \mid \vec{Z} \right)}{\Npop}\left(1 - \frac{\E\left(P_t \mid \vec{Z} \right)}{\Npop} \right) \\
    &= \V(Y_t \mid P_t = \E(P_t \mid \vec{Z}))
\end{align}
which is the required result.

>>>>

Starting with the variances, by the law of total variance:
\begin{align}
  &\var\left( Y_t \mid \vec{Z} \right) \\
    &= \var\left[\E\left( Y_t \mid \vec{Z}, \vec{P} \right) \mid \vec{Z} \right] + \E\left[\var\left( Y_t \mid \vec{Z}, \vec{P} \right) \mid \vec{Z} \right] \\
    &= \var\left( n_t \frac{P_t}{\Npop} \mid \vec{Z} \right) + \E\left( n_t \frac{P_t}{\Npop} \left(1 - \frac{P_t}{\Npop} \right) \mid \vec{Z} \right) &Y_t \mid \vec{P} \dist \Bin \\
    &= \left( \frac{n_t}{\Npop} \right)^2 \var\left(P_t \mid \vec{Z} \right) + \frac{n_t}{\Npop} \E\left(P_t \mid \vec{Z} \right)  - \frac{n_t}{\Npop^2} \E\left(P_t^2 \mid \vec{Z} \right) \\
    &= \left( \frac{n_t}{\Npop} \right)^2 \var\left(P_t \mid \vec{Z} \right) + \frac{n_t}{\Npop} \E\left(P_t \mid \vec{Z} \right)  - \frac{n_t}{\Npop^2} \left(\E\left(P_t \mid \vec{Z} \right) ^ 2 + \V(P_t \mid \vec{X})\right) \\
    &= \frac{n_t(n_t-1)}{\Npop} \var\left(P_t \mid \vec{Z} \right) + \frac{n_t}{\Npop} \E\left(P_t \mid \vec{Z} \right)  - \frac{n_t}{\Npop^2} \E\left(P_t \mid \vec{Z} \right) ^ 2.
    \label{inc-prev:eq:VY1}
\end{align}
In the CIS $n_t << \Npop$ because $\Npop$ is the population of England (around 56 million) while $n_t$ is the daily sample size (10s of thousands).
Therefore:
\begin{align}
    \frac{n_t(n_t - 1)}{\Npop^2} \var\left(P_t \mid \vec{Z} \right)
    &<< \frac{n_t}{\Npop} \var\left(P_t \mid \vec{Z} \right) \\
    &\leq \frac{n_t}{\Npop} \E\left(P_t \mid \vec{Z} \right) &\text{by \cref{inc-prev:eq:boundVPt}}.
\intertext{Therefore:}
\frac{n_t}{\Npop} \E\left(P_t \mid \vec{Z} \right) &\approx\frac{n_t(n_t-1)}{\Npop} \var\left(P_t \mid \vec{Z} \right) + \frac{n_t}{\Npop} \E\left(P_t \mid \vec{Z} \right).
\intertext{Substitute into \cref{inc-prev:eq:VY1}:}
  \var\left( Y_t \mid \vec{Z} \right)
    &\approx \frac{n_t}{\Npop} \E\left(P_t \mid \vec{Z} \right)  - \frac{n_t}{\Npop^2} \E\left(P_t \mid \vec{Z} \right) ^ 2 \\
    &= n_t \frac{\E\left(P_t \mid \vec{Z} \right)}{\Npop} \left(1 - \frac{\E\left(P_t \mid \vec{Z} \right)}{\Npop} \right) \\
    &= \V(Y_t \mid P_t = \E(P_t \mid \vec{Z}))
\end{align}
which is the required result.
